---
title: "Cleaning ovarian cancer"
author: "Kevin Wang"
date: "`r paste0('Initiated on 2020 Mar 4, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
   code_folding: hide
   fig_height: 10
   fig_width: 10
   toc: yes
   number_sections: true
   toc_depth: 3
   toc_float: yes
  editor_options: 
    chunk_output_type: console
---


# Summary

# Loading 

## Packages

```{r}
today = format(Sys.time(), "%Y_%b_%d")
cat("This file was compiled on", today)


suppressPackageStartupMessages({
  library(tidyverse)
  library(survival)
  library(survminer)
  library(ggsci)
  library(ggbeeswarm)
  
  library(curatedOvarianData)
})

theme_set(theme_classic(12) +
            theme(legend.position = "bottom"))
```



# Data

The Yoshihara 2012 CCR paper published a list of 126 genes. We will use this as the starting point to build our CPOP model. In the same paper, there are additional datasets used for validation. We will for now assume these data are of sufficient biological similarity to be applied. 

```{r japan_a}
data("GSE32062.GPL6480_eset")
japan_a = GSE32062.GPL6480_eset
```

```{r tothill}
data("GSE9891_eset")
tothill = GSE9891_eset
```

```{r japan_b}
data("GSE32063_eset")
japan_b = GSE32063_eset
```

```{r tcgarnaseq}
data("TCGA.RNASeqV2_eset")
selection = TCGA.RNASeqV2_eset$tumorstage %in% c(3, 4) & TCGA.RNASeqV2_eset$site_of_tumor_first_recurrence == "metastasis"
selection[is.na(selection)] = FALSE

tcgarnaseq = TCGA.RNASeqV2_eset[,selection]
```


<!-- ```{r tcgaarray} -->
<!-- data("TCGA_eset") -->

<!-- table(TCGA_eset$tumorstage, -->
<!--       TCGA_eset$summarystage) -->

<!-- table(TCGA_eset$tumorstage, -->
<!--       TCGA_eset$grade) -->

<!-- table(TCGA_eset$site_of_tumor_first_recurrence, -->
<!--       TCGA_eset$tumorstage) -->

<!-- tcgaarray = TCGA_eset[,TCGA_eset$tumorstage %in% c(3, 4) & TCGA_eset$site_of_tumor_first_recurrence == "metastasis"] -->
<!-- ``` -->


```{r bentink}
data("E.MTAB.386_eset")
bentink = E.MTAB.386_eset
```

```{r crijns}
data("GSE13876_eset")
crijns = GSE13876_eset

table(crijns$tumorstage,
      crijns$summarystage)

table(crijns$grade)

crijns = crijns[,crijns$grade %in% c(3, 4)]
```

```{r mok}
data("GSE18520_eset")
mok = GSE18520_eset
```

```{r}
# data("GSE49997_eset")
# pils = GSE49997_eset
```


```{r denkert}
# data("GSE14764_eset")
# denkert = GSE14764_eset
```

```{r yoshihara2010}
data("GSE17260_eset")
yoshihara2010 = GSE17260_eset
```


```{r}
data("GSE26712_eset")
bonome = GSE26712_eset
```


```{r}
# data("GSE30161_eset")
# ferriss = GSE30161_eset
```

```{r}
## Missing values exist
# data("GSE51088_eset")
# karlan = GSE51088_eset[,GSE51088_eset$tumorstage %in% c(3, 4)]
# karlan = karlan[(karlan %>% exprs %>% is.na %>% rowSums) == 0, ]
```

```{r}
data("GSE19829.GPL8300_eset")
konst = GSE19829.GPL8300_eset
```

```{r}
# data("PMID17290060_eset")
# dressman = PMID17290060_eset
```



```{r}
list_ovarian_eset = lst(
  japan_a, tothill, japan_b,
  tcgarnaseq, bonome, mok, yoshihara2010,
  bentink, crijns
  # dressman
  # konst, 
  )

list_ovarian_eset %>% 
  sapply(dim)
```

# Common genes between datasets
```{r}
raw_gene_list = purrr::map(list_ovarian_eset, rownames)

common_genes = Reduce(f = intersect, x = raw_gene_list)
length(common_genes)
```

# Survival samples
```{r}
ov_pdata = purrr::map(list_ovarian_eset, pData)


list_pdata =  list_ovarian_eset %>% 
  purrr::map(pData) %>% 
  purrr::map(tibble::rownames_to_column, var = "sample_id")

# list_pdata_filtered = list_pdata %>% 
#   purrr::map(.f = ~ .x %>% 
#                dplyr::filter(
#                  !(summarystage == "early" | summarygrade == "low" | grade %in% c(1,2) 
#                      # histological_type != "ser")
#                  )))
# sapply(list_pdata_filtered, nrow)

ov_surv_raw = purrr::map(
  .x = list_pdata, 
  .f = ~ data.frame(
    sample_id = .x$sample_id,
    time = .x$days_to_death %>% as.integer,
    dead = ifelse(.x$vital_status == "deceased", 1, 0)) %>%
    na.omit() %>% 
    dplyr::filter(time > 0, 
                  !is.nan(time),
                  !is.nan(dead))
)

ov_surv_raw %>% sapply(nrow)

ov_surv_y = ov_surv_raw %>% 
  purrr::map(~ .x %>% 
               dplyr::select(-sample_id)) %>% 
  purrr::map(~ Surv(time = .x$time, event = .x$dead))
```


# Reduced ovarian cancer
```{r}
ov_surv_exprs = purrr::map2(
  .x = list_ovarian_eset, 
  .y = ov_surv_raw,
  .f = ~ exprs(.x[common_genes,.y$sample_id])
)

ov_surv_exprs %>% sapply(dim)
```

# Binary samples 

```{r}
ov_surv_tbl = ov_surv_raw %>% 
  bind_rows(.id = "data_source")

ov_surv_tbl %>% 
  ggplot(aes(x = time)) +
  geom_histogram(aes(y = ..density..)) +
  facet_wrap(~data_source)


ov_surv_tbl %>% 
  ggplot(aes(x = time,
             y = ..density..,
             fill = data_source)) +
  geom_density(alpha = 0.2) +
  scale_fill_d3()

p = ov_surv_tbl %>% 
  ggplot(aes(x = time/365.25,
             y = ..density..,
             colour = factor(dead == 1))) +
  geom_density(alpha = 0.8, size = 2) +
  scale_colour_d3() +
  facet_wrap(~ data_source) +
  labs(x = "Time (years)") 

p

ggsave(p, filename = paste0("figures/ovarian_surv_distribution_", today, ".eps"))
```


# Save data
```{r}
save(
  ov_surv_exprs,
  ov_surv_raw,
  ov_pdata,
  ov_surv_y,
  file = "data/processed_data/ovarian_survival_28_Apr_2020.RData")
```

# Session Info 
```{r}
sessioninfo::session_info()
```
