---
title: "5 Checking microarray and nanostring consistency"
author: "Kevin Wang"
date: "`r paste0('Initiated on 2021 Jan 10, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 10
    fig_width: 10
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---


# Summary


# Loading 

## Packages
```{r}
today = format(Sys.time(), "%Y_%b_%d")
cat("This file was compiled on", today)


suppressPackageStartupMessages({
  library(tidyverse)
  
  library(survminer)
  library(survival)
  
  library(limma)
  library(ClassifyR)
  library(patchwork)
})

theme_set(
  theme_classic(12) +
    theme(legend.position = "bottom",
          axis.text = element_text(colour = "black"),
          axis.ticks = element_line(colour = "black")))
```

## Data
```{r}
file_version = "2020_Apr_25"
load(paste0("../data/processed_data/Melanoma4_binary_", file_version, ".RData"))
```


# Microarray - DE analysis
```{r}
array_rfs_class = list_binary_lassoy$array
design_rfs_array = model.matrix(~0+array_rfs_class)
design_rfs_array_contr = makeContrasts(
  array_rfs_classPoor - array_rfs_classGood,
  levels = design_rfs_array)

rfs_array_exprs_data = list_raw_data_binary$array

fit_rfs_array = lmFit(
  rfs_array_exprs_data, 
  design = design_rfs_array)


fit_rfs_array = contrasts.fit(fit_rfs_array, design_rfs_array_contr)
fit_rfs_array = eBayes(fit_rfs_array)
fit_rfs_array_table = topTable(fit_rfs_array, 
                               adjust.method = "fdr", 
                               number = "all") %>% 
  tibble::rownames_to_column("genename") %>% 
  tibble::as_tibble()

fit_rfs_array_table
```

# NanoString - DE analysis

```{r}
ns_rfs_class = list_binary_lassoy$ns
design_rfs_ns = model.matrix(~0+ns_rfs_class)
design_rfs_ns_contr = makeContrasts(
  ns_rfs_classPoor - ns_rfs_classGood,
  levels = design_rfs_ns)

rfs_ns_exprs_data = list_raw_data_binary$ns

fit_rfs_ns = lmFit(
  rfs_ns_exprs_data, 
  design = design_rfs_ns)


fit_rfs_ns = contrasts.fit(fit_rfs_ns, design_rfs_ns_contr)
fit_rfs_ns = eBayes(fit_rfs_ns)
fit_rfs_ns_table = topTable(fit_rfs_ns, 
                               adjust.method = "fdr", 
                               number = "all") %>% 
  tibble::rownames_to_column("genename") %>% 
  tibble::as_tibble()

fit_rfs_ns_table
```

# Combining microarray and NS

## Log-fold

```{r}
combine_limma_table_long = 
  bind_rows(fit_rfs_array_table %>% mutate(type = "microarray"), 
            fit_rfs_ns_table %>% mutate(type = "ns"))

## Join two tables by gene column only
combine_limma_table_wide = fit_rfs_array_table %>% 
  dplyr::left_join(fit_rfs_ns_table, 
                   by = "genename", suffix = c("_array", "_ns"))
```

```{r}
compare_logFC = combine_limma_table_wide %>% 
  ggplot(aes(x = logFC_ns,
             y = logFC_array)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, colour = "red") +
  labs(x = "logFC on NanoString",
       y = "logFC on Microarray",
       title = "Compare logFC between the two platforms")

compare_logFC

cat("The correlation between the two sets of logFC is ", 
    cor(combine_limma_table_wide$logFC_ns, combine_limma_table_wide$logFC_array))

ggsave(filename = paste0("../figures/compare_logFC_", file_version, ".eps"),
       plot = compare_logFC,
       width = 6, height = 6)
```


## Compare t-statistics
```{r}
compare_t = combine_limma_table_wide %>% 
  ggplot(aes(x = t_ns,
             y = t_array)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, colour = "red") +
  labs(x = "t-statistics on NanoString",
       y = "t-statistics on Microarray",
       title = "Compare t-statistics between the two platforms")

compare_t

cat("The correlation between the two sets of t-statistics is ", 
    cor(combine_limma_table_wide$t_ns, combine_limma_table_wide$t_array))

ggsave(filename = paste0("../figures/compare_tstat_", file_version, ".eps"),
       plot = compare_t,
       width = 6, height = 6)
```

# Full JID79 sample data (MIA-cohort)

```{r}
raw2018_sample_data = readxl::read_excel("../data/raw_data/JID_79_FUP&Rx_11Apr2018.xlsx") %>% 
  janitor::clean_names()

clean2018_sample_data = raw2018_sample_data %>% 
  dplyr::mutate(
    tbid = paste0("TB", tumour_id),
    
    date_first_recurrence_post_banking_corrected = case_when(
      complete.cases(date_first_recurrence_post_banking) ~ date_first_recurrence_post_banking,
      person_fu_status == "Alive NSR" ~ person_date_last_fu_death
    ),
    
    date_first_distant_met_post_banking_corrected = case_when(
      complete.cases(date_first_distant_met_post_banking) ~
        date_first_distant_met_post_banking,
      person_fu_status == "Alive NSR" ~ person_date_last_fu_death
    ),
    RFS = as.integer(
      difftime(date_first_recurrence_post_banking_corrected,
               tumour_date_banked,
               units = "days")/30),
    DMS = as.integer(
      difftime(date_first_distant_met_post_banking_corrected,
               tumour_date_banked,
               units = "days")/30),
    OS = as.integer(difftime(person_date_last_fu_death, 
                             tumour_date_banked,
                             units = "days")/30),
    dead = ifelse(str_detect(person_fu_status, "Alive"), 0, 1)
  )
```

# KM survival plots

## RFS
```{r}
rfs_surv = survfit(Surv(RFS, dead) ~ 1, data = clean2018_sample_data)

surv_median(rfs_surv)

ggsurvplot(rfs_surv, data = clean2018_sample_data,
           surv.median.line = "hv", # Add medians survival
           legend.title = "RFS",
           risk.table = TRUE,
           tables.height = 0.2,
           tables.theme = theme_cleantable())
```



## DMS
```{r}
dms_surv = survfit(Surv(DMS, dead) ~ 1, data = clean2018_sample_data)

surv_median(dms_surv)

ggsurvplot(dms_surv, data = clean2018_sample_data,
           surv.median.line = "hv", # Add medians survival
           legend.title = "DMS",
           risk.table = TRUE,
           tables.height = 0.2,
           tables.theme = theme_cleantable())
```



## OS
```{r}
os_surv = survfit(Surv(OS, dead) ~ 1, data = clean2018_sample_data)

surv_median(os_surv)

ggsurvplot(os_surv, data = clean2018_sample_data,
           surv.median.line = "hv", # Add medians survival
           legend.title = "os",
           risk.table = TRUE,
           tables.height = 0.2,
           tables.theme = theme_cleantable())
```

# Loading full microarray and nanostring data

## Microarray - 79 samples identical to the JID79 cohort

```{r}
file_version = "2020_Jan_12"
load(paste0("../data/processed_data/Chip1to12_144samples_matched_",
            file_version, ".RData"))
```

```{r, fig.height=6,fig.width=6}
gplots::venn(list(
  microarray_samples = colnames(array_exprs_data),
  cleaned2018_samples = clean2018_sample_data$tbid
))
```

## NanoString - 144 samples (incl replicates)

The NanoString platform accommodates 12 samples per chip. And a total of 12 chips data were generated in 2018. These samples include some replicates (hence the overlap is more than 79 samples). The NS samples, is a super-set of the binary classification of OS/RFS when the microarray analysis was performed in 2012. 

```{r, fig.height=6,fig.width=6}
gplots::venn(list(
  ns_144_samples = ns_cleaned_sample_data$sample_id,
  cleaned2018_samples = clean2018_sample_data$tbid
))

gplots::venn(list(
  ns_unique_samples = ns_cleaned_sample_data$sample_id %>% unique,
  cleaned2018_samples = clean2018_sample_data$tbid
))
```


# Creating sample class using RFS

## Microarray
```{r}
array_2018sample_data = clean2018_sample_data %>% 
  dplyr::mutate(
    rfs_class = case_when(
      (RFS >= 4*12 & person_fu_status == "Alive NSR") ~ "Good",
      (RFS <= 1*12 & person_fu_status == "Dead, melanoma") ~ "Poor",
      TRUE ~ NA_character_
    ) %>% as.factor(),
    os_class = case_when(
      (OS >= 4*12 & person_fu_status == "Alive NSR") ~ "Good",
      (OS <= 1*12 & person_fu_status == "Dead, melanoma") ~ "Poor",
      TRUE ~ NA_character_
    ) %>% as.factor()
  )

with(array_2018sample_data, 
     table(os_class, rfs_class, useNA = "always"))

os_vs_rfs = array_2018sample_data %>% 
  ggplot(aes(
    x = OS, y = RFS,
    colour = ifelse(complete.cases(os_class), as.character(os_class), "Neither"),
    shape = ifelse(complete.cases(rfs_class), as.character(rfs_class), "Neither"))) +
  geom_point(size = 3, stroke = 1.2) +
  scale_shape_discrete(solid = FALSE) +
  labs(colour = "OS classification",
       shape = "RFS classification",
       title = "Comparing samples with OS and non-missing RFS")

os_vs_rfs
```

## NanoString (unique sample with top count)
```{r}
ns_cleaned_sample_data_unique = ns_cleaned_sample_data %>% 
  dplyr::mutate(
    total_counts = colSums(ns_cleaned_exprs_data)
  ) %>% 
  group_by(sample_id) %>% 
  dplyr::arrange(-total_counts) %>% 
  dplyr::distinct(sample_id, .keep_all = TRUE) %>% 
  ungroup()


ns_2018sample_data_unique = clean2018_sample_data %>% 
  dplyr::left_join(
    ns_cleaned_sample_data_unique, 
    by = c("tbid" = "sample_id")) %>% 
  dplyr::mutate(
    rfs_class = case_when(
      (RFS >= 4*12 & person_fu_status == "Alive NSR") ~ "Good",
      (RFS <= 1*12 & person_fu_status == "Dead, melanoma") ~ "Poor",
      TRUE ~ NA_character_) %>% as.factor(),
    os_class = case_when(
      (OS >= 4*12 & person_fu_status == "Alive NSR") ~ "Good",
      (OS <= 1*12 & person_fu_status == "Dead, melanoma") ~ "Poor",
      TRUE ~ NA_character_
    ) %>% as.factor())

with(ns_2018sample_data_unique, 
     table(os_class, rfs_class, useNA = "always"))
```


# Correlate all genes 
```{r}
rfs_ns_exprs_data %>% dim
rfs_array_exprs_data %>% dim

common_genes = intersect(
  rownames(rfs_ns_exprs_data),
  rownames(rfs_array_exprs_data))


# rfs_ns_exprs_data[common_genes, ], rfs_array_exprs_data[common_genes,]

common_genes_corr = tibble(
  common_genes, 
  ns = purrr::map(.x = common_genes, .f = ~ rfs_ns_exprs_data[.x, ]),
  array = purrr::map(.x = common_genes, .f = ~ rfs_array_exprs_data[.x, ]),
  corr = purrr::map_dbl(.x = common_genes,
           .f = ~ cor(rfs_ns_exprs_data[.x, ], rfs_array_exprs_data[.x,])))


common_genes_corr_boxplot = common_genes_corr %>% 
  ggplot(aes(x = "1", y = corr)) +
  geom_boxplot() +
  labs(x = "", 
       y = "Correlation")



common_genes_corr_top6 = common_genes_corr %>% 
  dplyr::filter(rank(-corr) <= 6) %>% 
  unnest(c(ns, array)) %>% 
  ggplot(aes(x = ns, y = array)) +
  geom_point() +
  ggpubr::stat_cor() +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  facet_wrap(~common_genes, scales = "free") +
  labs(x = "NanoString expression", 
       y = "Microarray expression")



ggsave(filename = paste0("../figures/corr_ns_array_genes_", file_version, ".eps"),
       plot = common_genes_corr_top6 + 
  common_genes_corr_boxplot + 
  patchwork::plot_layout(widths = c(3, 1)),
       width = 10, height = 6)
```


# Curating RFS samples before saving

## Subsetting genes

```{r}
reduced_genes = intersect(
  rfs_array_exprs_data %>% rownames,
  rfs_ns_exprs_data %>% rownames  
) %>% sort

rfs_array_exprs_data_reduced = rfs_array_exprs_data[reduced_genes, ]
rfs_ns_exprs_data_reduced = rfs_ns_exprs_data[reduced_genes, ]
```


## Subsetting samples 

```{r}
array_2018sample_data_filtered = array_2018sample_data %>% 
  dplyr::filter(complete.cases(rfs_class))
rfs_array_sample_data_reduced = array_2018sample_data_filtered

ns_2018sample_data_filtered = ns_2018sample_data_unique %>% 
  dplyr::filter(complete.cases(rfs_class))
rfs_ns_sample_data_reduced = ns_2018sample_data_filtered

colnames(rfs_ns_exprs_data_reduced) = rfs_ns_sample_data_reduced$tbid
```


```{r}
stopifnot(identical(
  rfs_array_exprs_data_reduced %>% dimnames,
  rfs_ns_exprs_data_reduced %>% dimnames
))
```

# ClassifyR on the classifications

```{r, eval = FALSE}
nPerm = 100
nCores = 5

resubstituteParams = ResubstituteParams(
  nFeatures = c(20, 50, 100),
  performanceType = "balanced error",
  better = "lower")

crParams = list(
  SelectParams(limmaSelection, 
               "Moderated t Statistic",
               resubstituteParams = resubstituteParams),
  TrainParams(SVMtrainInterface),
  PredictParams(SVMpredictInterface))
```

## Array + OS
```{r, eval = FALSE}
array_exprs_data %>% dim
table(array_2018sample_data$rfs_class, useNA = "always")

array_os_sample_data = array_2018sample_data %>% 
  dplyr::filter(complete.cases(os_class))

array_os_exprs_data = array_exprs_data[unique(rownames(array_exprs_data)),array_os_sample_data$tbid]

array_os_class_results <- runTests(
  array_os_exprs_data,
  array_os_sample_data$os_class,
  datasetName = "array_os_classifyR",
  classificationName = "array_os", 
  permutePartition = "fold",
  permutations = nPerm, folds = 5,
  verbose = 3, seed = 2018,
  parallelParams = MulticoreParam(workers = nCores),
  params = crParams
)
```

## Array + RFS


```{r, eval = FALSE}
array_exprs_data %>% dim
table(array_2018sample_data$rfs_class, useNA = "always")

array_rfs_sample_data = array_2018sample_data %>% 
  dplyr::filter(complete.cases(rfs_class))

array_rfs_exprs_data = array_exprs_data[unique(rownames(array_exprs_data)),array_rfs_sample_data$tbid]

array_rfs_class_results <- runTests(
  array_rfs_exprs_data,
  array_rfs_sample_data$rfs_class,
  datasetName = "array_rfs_classifyR",
  classificationName = "array_rfs", 
  permutePartition = "fold",
  permutations = nPerm, folds = 5,
  verbose = 3, seed = 2018,
  parallelParams = MulticoreParam(workers = nCores),
  params = crParams
)
```


## NS + OS

```{r, eval = FALSE}
ns_os_sample_data = ns_2018sample_data_unique %>% 
  dplyr::filter(complete.cases(os_class),
                file_name %in% colnames(ns_cleaned_exprs_data))

ns_os_exprs_data = log2(ns_cleaned_exprs_data[,ns_os_sample_data$file_name] + 1L)


ns_os_class_results <- runTests(
  ns_os_exprs_data,
  ns_os_sample_data$os_class,
  datasetName = "ns_os_classifyR",
  classificationName = "ns_os", 
  permutePartition = "fold",
  permutations = nPerm, folds = 5,
  verbose = 3, seed = 2018,
  parallelParams = MulticoreParam(workers = nCores),
  params = crParams
)

```


## NS + RFS

```{r, eval = FALSE}
ns_rfs_sample_data = ns_2018sample_data_unique %>% 
  dplyr::filter(complete.cases(rfs_class),
                file_name %in% colnames(ns_cleaned_exprs_data))

ns_rfs_exprs_data = log2(ns_cleaned_exprs_data[,ns_rfs_sample_data$file_name] + 1L)


ns_rfs_class_results <- runTests(
  ns_rfs_exprs_data,
  ns_rfs_sample_data$rfs_class,
  datasetName = "ns_rfs_classifyR",
  classificationName = "ns_rfs", 
  permutePartition = "fold",
  permutations = nPerm, folds = 5,
  verbose = 3, seed = 2018,
  parallelParams = MulticoreParam(workers = nCores),
  params = crParams
)
```

## Comparing results 

```{r}
# list_classifyr_results = list(
#   "array_RFS" = array_rfs_class_results,
#   "array_OS" = array_os_class_results,
#   "NS_OS" = ns_os_class_results,
#   "NS_RFS" = ns_rfs_class_results)
# saveRDS(list_classifyr_results, file = "../data/processed_data/list_array_ns_classifyR.rds")
```

```{r}
list_classifyr_results = readRDS(file = "../data/processed_data/list_array_ns_classifyR.rds")

list_perform = list_classifyr_results %>% 
  purrr::map(calcCVperformance, "accuracy") %>% 
  purrr::map(~unlist(.x@performance))

perform_tbl = list_perform %>% 
  reshape2::melt() %>% 
  transmute(accuracy = value,
            type = L1) %>% 
  as_tibble()

q1 = function(x){quantile(x, 0.25)}
q3 = function(x){quantile(x, 0.75)}

perform_plot = perform_tbl %>% 
  ggplot(aes(x = type, y = accuracy)) +
  geom_boxplot() +
  labs(x = "Data/Prognosis definition", 
       y = "Accuracy",
       title = "Accuracy across MIA data and prognosis class definitions")


perform_plot

ggsave(filename = paste0("../figures/perform_plot_", file_version, ".eps"),
       plot = perform_plot,
       width = 7, height = 6)


pairwise.wilcox.test(x = perform_tbl$accuracy, 
                     g = perform_tbl$type, p.adjust.method = "fdr")
```

# Session Info 
```{r}
sessioninfo::session_info()
```
