---
title: "8 CPOP vs Combat normalised Lasso predictions"
author: "Kevin Wang"
date: "`r paste0('Initiated on 2022 Jan 10, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 10
    fig_width: 10
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---


# Summary

+ When using Lasso on the MIA-NS or MIA-Microarray data, the features selected are not identical, with many features being selected by only one dataset but not the other. 

+ This does not improve when we look at the correlation of prediction value.


```{r}
today = format(Sys.time(), "%Y_%b_%d")
cat("This file was compiled on", today)


suppressPackageStartupMessages({
  library(tidyverse)
  library(CPOP)
  library(glmnet)
  library(survival)
  library(survminer)
  library(patchwork)
  library(sva)
  library(fairness)
  library(yardstick)
})

theme_set(theme_classic(16) +
            theme(legend.position = "bottom", 
                  axis.text = element_text(colour = "black"),
                  axis.ticks = element_line(colour = "black")))
```


## data

```{r}
file_version = "2020_Apr_25"
load(paste0(
  "../data/processed_data/Melanoma4_binary_",
  file_version, ".RData"
))


load(paste0(
  "../data/processed_data/Melanoma4_survival_",
  file_version, ".RData"
))
```


```{r}
list_lratio_binary = list_raw_data_binary %>%
  purrr::map(t) %>%
  purrr::map(CPOP::pairwise_col_diff)

list_lratio_survival = list_raw_data %>%
  purrr::map(t) %>%
  purrr::map(CPOP::pairwise_col_diff)
```

## Customised functions
```{r}
join_tcga_sample_data = function(dat){
  dat %>% 
    left_join(tcga_sample_data, by = c("samples" = "sample_id")) %>% 
    dplyr::filter(naive_tumor_stage == "stageIII", !str_detect(samples, "TCGA-EE")) %>% 
    dplyr::mutate(time = time/365) %>% 
    dplyr::mutate(
    tcga_median_class = case_when(
      status == 0 & time > median(time) ~ "Good",
      status == 1 & time < median(time) ~ "Poor",
      TRUE ~ NA_character_))
}

join_sweden_sample_data = function(dat){
  dat %>% 
    left_join(sweden_sample_data, by = c("samples" = "sample_id")) %>% 
    dplyr::mutate(time = dss/365, status = dss_dead) %>% 
    dplyr::mutate(
    sweden_median_class = case_when(
      status == 0 & time > median(time) ~ "Good",
      status == 1 & time < median(time) ~ "Poor",
      TRUE ~ NA_character_))
}

eval_prediction_df = function(dat){
  cat("Predicted classes: \n")
  table(dat$pred_class)
  
  cat("KM-plots: \n")
  print(
    survminer::ggsurvplot(
      fit = survfit(Surv(time, status) ~ pred_class, 
                    data = dat), 
      data = dat, 
      break.time.by = 5,
      pval = TRUE, risk.table = "nrisk_cumevents"))
  
  surv_obj = coxph(Surv(time, status) ~ pred_class + age + gender, 
          data = dat)
  
  cat("Cox regression summary: \n")
  summary(surv_obj)
  broom::glance(surv_obj, data = dat) %>% glimpse
  
  # equal_odds(
  #   data  = dat %>% 
  #     dplyr::mutate(
  #       surv_median_class = case_when(
  #         status == 0 & time > median(time) ~ "Good",
  #         status == 1 & time < median(time) ~ "Poor",
  #         TRUE ~ NA_character_)) %>% 
  #     dplyr::filter(complete.cases(surv_median_class)),
  #   outcome = 'surv_median_class',
  #   probs   = 'pred_prob',
  #   group   = 'gender',
  #   cutoff  = 0.5)
}
```


<!-- # Survival times distributional comparison  -->


<!-- ```{r} -->
<!-- list_surv_times = lst( -->
<!--   mia = tibble( -->
<!--     time = rfs_array_sample_data_reduced$RFS,  -->
<!--     dead = rfs_array_sample_data_reduced$dead),  -->
<!--   tcga = tibble( -->
<!--     time = tcga_sample_data$time/365,  -->
<!--     dead = tcga_sample_data$status), -->
<!--   sweden = tibble( -->
<!--     time = sweden_sample_data$dss/365,  -->
<!--     dead = sweden_sample_data$dss_dead)) -->

<!-- surv_times_tbl = list_surv_times %>% bind_rows(.id = "data_source") -->

<!-- list_survival_lassoy = list( -->
<!--   ns = Surv(rfs_ns_sample_data_reduced$RFS + 1, -->
<!--             rfs_ns_sample_data_reduced$dead), -->
<!--   array = Surv(rfs_array_sample_data_reduced$RFS + 1, -->
<!--                rfs_array_sample_data_reduced$dead), -->
<!--   tcga = Surv(tcga_sample_data$time, -->
<!--                tcga_sample_data$status), -->
<!--   sweden = Surv(sweden_sample_data$dss, -->
<!--                sweden_sample_data$dss_dead)) -->
<!-- ``` -->

# Compute a single CPOP model from MIA-NS and MIA-Microarray (saved)
```{r}
# set.seed(12)
# 
# mia_cpop_binary = CPOP::cpop_model(
#   z1 = list_lratio_binary$ns,
#   z2 = list_lratio_binary$array,
#   y1 = list_binary_lassoy$ns,
#   y2 = list_binary_lassoy$array,
#   family = "binomial",
#   alpha = 0.1,
#   n_features = 20)
# save(mia_cpop_binary, file = "../data/processed_data/shiny_models_2021_Nov_02.RData")
load(file = "../data/processed_data/shiny_models_2021_Nov_02.RData")
```

## CPOP binary evaluations
```{r}
mia_cpop_pred_tcga = predict_cpop(mia_cpop_binary, newz = list_lratio_binary$tcga) %>% 
  join_tcga_sample_data() %>% 
  dplyr::rename(
    pred_prob = cpop_model_avg_prob,
    pred_class = cpop_model_avg_class)
  # dplyr::filter(complete.cases(tcga_median_class))

# eval_prediction_df(mia_cpop_pred_tcga)
```


```{r}
mia_cpop_pred_tcga %>% 
  ggplot(aes(x = CPOP::expit(cpop_model1),
             y = CPOP::expit(cpop_model2))) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(colour = "red") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1))

cat("Correlation: ", 
    cor(CPOP::expit(mia_cpop_pred_tcga$cpop_model1), 
        CPOP::expit(mia_cpop_pred_tcga$cpop_model2)), "\n")

cat("Concordance (yardstick): ", 
    yardstick::ccc_vec(mia_cpop_pred_tcga$cpop_model1, 
                       mia_cpop_pred_tcga$cpop_model2), "\n")

cat("Ideality of correlation: ", 
    yardstick::iic_vec(mia_cpop_pred_tcga$cpop_model1, 
                       mia_cpop_pred_tcga$cpop_model2), "\n")
```


## CPOP survival evaluation
```{r}
mia_cpop_pred_tcga_surv = predict_cpop(mia_cpop_binary, newz = list_lratio_survival$tcga) %>% 
  join_tcga_sample_data() %>% 
  dplyr::rename(
    pred_prob = cpop_model_avg_prob,
    pred_class = cpop_model_avg_class)

eval_prediction_df(mia_cpop_pred_tcga_surv)
```



# Training Lasso (elastic net) on MIA-NS and MIA-Microarray and predicting on TCGA

Notice now one of the data (MIA-NS) tends to select more features, but a majority of these features were not agreed by MIA-Microarray. This disagreement is a fundamental motivation for developing CPOP.

Furthermore, when breaking up this data by the sign of the constructed Elastic Net model coefficients, we see the disagreed features are also not balanced in sign. This has implications on the final predicted values. Which we will demonstrate in subsections.

```{r}
set.seed(2022)

mia_lasso_binary = CPOP::naive_glmnet(
  x1 = list_raw_data_binary$ns %>% t,
  x2 = list_raw_data_binary$array %>% t,
  y1 = list_binary_lassoy$ns,
  y2 = list_binary_lassoy$array,
  family = "binomial",
  alpha = 0.1)

# mia_lasso_binary$coef_tbl[[1]] %>%
#   dplyr::filter(coef1 != 0 | coef2 != 0, coef_name != "(Intercept)") %>%
#   ggplot(aes(x = coef1, y = coef2, label = coef_name)) +
#   geom_point() +
#   geom_abline(colour = "red")

mia_lasso_binary_selected_tbl = mia_lasso_binary$coef_tbl[[1]] %>%
  dplyr::mutate(selected_ns = (sign(coef1) != 0),
                selected_array = (sign(coef2) != 0))

ms = yardstick::metric_set(f_meas, bal_accuracy)

mia_lasso_binary_selected_tbl %>% 
  dplyr::mutate(selected_ns = factor(selected_ns),
                selected_array = factor(selected_array)) %>% 
  ms(truth = selected_ns, estimate = selected_array)

mia_lasso_binary_selected_tbl %>% 
  group_by(selected_ns, selected_array) %>% 
  tally() %>% 
  ungroup() %>% 
  ggplot(aes(x = selected_ns, y = selected_array, label = n)) +
  geom_tile(colour = "grey20", fill = "grey80") +
  geom_text(size = 5) +
  scale_fill_distiller(palette = "Blues", direction = 1)

mia_lasso_binary$coef_tbl[[1]] %>%
  group_by(sign_ns = sign(coef1), sign_array = sign(coef2)) %>%
  tally() %>%
  ungroup() %>%
  tidyr::complete(sign_ns, sign_array, fill = list(n = 0)) %>%
  ggplot(aes(x = sign_ns, y = sign_array,
             label = n)) +
  geom_tile(colour = "grey20", fill = "grey80") +
  geom_text(size = 5) +
  scale_fill_distiller(palette = "Blues", direction = 1) +
  labs(x = "Sign of MIA-NanoString Lasso model coef.",
       y = "Sign of MIA-Microarray Lasso model coef.")
```


## Naive Lasso survival evaluation

NanoString:

```{r}
ns_lasso_pred_tcga_surv = tibble(
    samples = rownames(list_lratio_survival$tcga),
    pred_prob = predict(mia_lasso_binary$glmnet1, 
                        newx = list_lratio_survival$tcga, 
                        type = "response", s = "lambda.min") %>% as.vector(),
    pred_class = predict(mia_lasso_binary$glmnet1, 
                         newx = list_lratio_survival$tcga, 
                         type = "class", s = "lambda.min") %>% as.vector()) %>%
  join_tcga_sample_data()

eval_prediction_df(ns_lasso_pred_tcga_surv)
```

Array

```{r}
array_lasso_pred_tcga_surv = tibble(
    samples = rownames(list_lratio_survival$tcga),
    pred_prob = predict(mia_lasso_binary$glmnet2, 
                        newx = list_lratio_survival$tcga, 
                        type = "response", s = "lambda.min") %>% as.vector(),
    pred_class = predict(mia_lasso_binary$glmnet2, 
                         newx = list_lratio_survival$tcga, 
                         type = "class", s = "lambda.min") %>% as.vector()) %>%
  join_tcga_sample_data()

eval_prediction_df(array_lasso_pred_tcga_surv)
```

# Overlapping the top k features via t-stat
```{r}
library(genefilter)
ns_colttest = genefilter::colttests(x = list_lratio_binary$ns, fac = list_binary_lassoy$ns) %>% 
  tibble::rownames_to_column("lratio_name") %>% as_tibble() %>% 
  dplyr::mutate(abs_stat = abs(statistic),
                rank_stat = rank(-abs_stat))
array_colttest = genefilter::colttests(x = list_lratio_binary$array, fac = list_binary_lassoy$array) %>% 
  tibble::rownames_to_column("lratio_name") %>% as_tibble() %>% 
  dplyr::mutate(abs_stat = abs(statistic),
                rank_stat = rank(-abs_stat))

tcga_colttest = genefilter::colttests(x = list_lratio_binary$tcga, fac = factor(list_binary_lassoy$tcga)) %>% 
  tibble::rownames_to_column("lratio_name") %>% as_tibble() %>% 
  dplyr::mutate(abs_stat = abs(statistic),
                rank_stat = rank(-abs_stat))

top_k_tbl = tibble(
  k = 1:500,
  `(NS, Microarray)` = purrr::map_int(
    .x = k,
    .f = ~ {
      length(intersect(
        ns_colttest %>% dplyr::filter(rank_stat <= .x) %>% dplyr::pull(lratio_name),
        array_colttest %>% dplyr::filter(rank_stat <= .x) %>% dplyr::pull(lratio_name)))
    }),
  `(NS, TCGA)` = purrr::map_int(
    .x = k,
    .f = ~ {
      length(intersect(
        ns_colttest %>% dplyr::filter(rank_stat <= .x) %>% dplyr::pull(lratio_name),
        tcga_colttest %>% dplyr::filter(rank_stat <= .x) %>% dplyr::pull(lratio_name)))
    }),
  `(Microarray, TCGA)` = purrr::map_int(
    .x = k,
    .f = ~ {
      length(intersect(
        array_colttest %>% dplyr::filter(rank_stat <= .x) %>% dplyr::pull(lratio_name),
        tcga_colttest %>% dplyr::filter(rank_stat <= .x) %>% dplyr::pull(lratio_name)))
    })
)

top_k_tbl %>%
  pivot_longer(cols = -c("k"), 
               names_to = "type", 
               values_to = "count") %>% 
  dplyr::mutate(proportion = count/k) %>% 
  ggplot(aes(x = k, y = proportion,
             colour = type)) +
  geom_line(size = 1.2) +
  scale_colour_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Proportion of overlap in the top k features",
       x = "k",
       colour = "Pairings")
```



<!-- ## Using `sva::ComBat` on NS+TCGA and predict on TCGA -->

<!-- ```{r} -->
<!-- ns_tcga_combat = sva::ComBat( -->
<!--   dat = rbind(list_lratio_binary$ns, list_lratio_binary$tcga) %>% t, -->
<!--   batch = c(rep("NS", nrow(list_lratio_binary$ns)), rep("TCGA", nrow(list_lratio_binary$tcga))), -->
<!--   mod = NULL, par.prior = TRUE, mean.only = FALSE) -->

<!-- ns_combat = ns_tcga_combat[,rownames(list_lratio_binary$ns)] %>% t -->
<!-- tcga_combat = ns_tcga_combat[,rownames(list_lratio_binary$tcga)] %>% t -->
<!-- ``` -->

<!-- ```{r} -->
<!-- set.seed(2022) -->

<!-- ns_combat_lasso = glmnet::cv.glmnet( -->
<!--   x = ns_combat, -->
<!--   y = list_binary_lassoy$ns, -->
<!--   family = "binomial", alpha = 0.1) -->

<!-- ns_combat_lasso_pred_tcga = tibble( -->
<!--     samples = rownames(list_lratio_binary$tcga), -->
<!--     pred_prob = predict(ns_combat_lasso,  -->
<!--                         newx = list_lratio_binary$tcga, type = "response", s = "lambda.min") %>% as.vector(), -->
<!--     pred_class = predict(ns_combat_lasso,  -->
<!--                          newx = list_lratio_binary$tcga, type = "class", s = "lambda.min") %>% as.vector()) %>% -->
<!--   join_tcga_sample_data() -->
<!-- ``` -->

<!-- ## Using `sva::ComBat` on Microarray+TCGA and predict on TCGA -->

<!-- ```{r} -->
<!-- array_tcga_combat = sva::ComBat( -->
<!--   dat = rbind(list_lratio_binary$array, list_lratio_binary$tcga) %>% t, -->
<!--   batch = c(rep("array", nrow(list_lratio_binary$array)), rep("TCGA", nrow(list_lratio_binary$tcga))), -->
<!--   mod = NULL, par.prior = TRUE, mean.only = FALSE) -->

<!-- array_combat = array_tcga_combat[,rownames(list_lratio_binary$array)] %>% t -->
<!-- tcga_combat = array_tcga_combat[,rownames(list_lratio_binary$tcga)] %>% t -->
<!-- ``` -->

<!-- ```{r} -->
<!-- set.seed(2022) -->

<!-- array_combat_lasso = glmnet::cv.glmnet( -->
<!--   x = array_combat, -->
<!--   y = list_binary_lassoy$array, -->
<!--   family = "binomial", alpha = 0.1) -->

<!-- array_combat_lasso_pred_tcga = tibble( -->
<!--     samples = rownames(list_lratio_binary$tcga), -->
<!--     pred_prob = predict(array_combat_lasso,  -->
<!--                         newx = list_lratio_binary$tcga, type = "response", s = "lambda.min") %>% as.vector(), -->
<!--     pred_class = predict(array_combat_lasso,  -->
<!--                          newx = list_lratio_binary$tcga, type = "class", s = "lambda.min") %>% as.vector()) %>% -->
<!--   join_tcga_sample_data() -->
<!-- ``` -->

<!-- ## Comparing prediction scores for `ComBat` samples -->

<!-- ```{r} -->
<!-- combine_combat_pred_tcga = inner_join( -->
<!--     x = ns_combat_lasso_pred_tcga %>% dplyr::select(samples, pred_prob, pred_class), -->
<!--     y = array_combat_lasso_pred_tcga %>% dplyr::select(samples, pred_prob, pred_class), -->
<!--     by = "samples", suffix = c("_ns", "_array")) -->

<!-- combine_combat_pred_tcga %>% -->
<!--   ggplot(aes(x = pred_prob_ns, y = pred_prob_array)) + -->
<!--   geom_point() + -->
<!--   geom_smooth(method = "lm") + -->
<!--   geom_abline(colour = "red") + -->
<!--   scale_x_continuous(limits = c(0, 1)) + -->
<!--   scale_y_continuous(limits = c(0, 1)) -->

<!-- cat("Correlation: ", cor(combine_combat_pred_tcga$pred_prob_ns,  -->
<!--                          combine_combat_pred_tcga$pred_prob_array), "\n") -->
<!-- ``` -->

# Normalise NS, Array, TCGA and predict on TCGA

```{r,fig.width=14}
mia_tcga_mat = rbind(
  list_lratio_binary$ns, 
  list_lratio_binary$array, 
  list_lratio_survival$tcga) %>% t
mia_tcga_colnames = c(
  paste0(rownames(list_lratio_binary$ns), "_ns"),
  paste0(rownames(list_lratio_binary$array), "_array"),
  paste0(rownames(list_lratio_survival$tcga), "_tcga"))
colnames(mia_tcga_mat) = mia_tcga_colnames

mia_tcga_combat = sva::ComBat(
  dat = mia_tcga_mat,
  batch = c(rep("ns", nrow(list_lratio_binary$ns)), 
            rep("array", nrow(list_lratio_binary$array)),
            rep("tcga", nrow(list_lratio_survival$tcga))))

ns_combat = mia_tcga_combat[,str_detect(mia_tcga_colnames, "_ns$")] %>% t
array_combat = mia_tcga_combat[,str_detect(mia_tcga_colnames, "_array$")] %>% t
tcga_combat = mia_tcga_combat[,!str_detect(mia_tcga_colnames, "_ns$|_array")] %>% t

mia_tcga_mat[1:5,1:5]
mia_tcga_combat[1:5,1:5]

mia_tcga_mat %>% 
  CPOP::boxplot_tbl(2) %>% 
  tidyr::separate(col = "object", into = c("sample_ID", "platform"), 
                  sep = "_", remove = FALSE) %>% 
  ggplot(aes(x = object, y = means,
             ymin = q1, ymax = q3,
             colour = platform)) +
  geom_point() +
  geom_linerange() +
  scale_colour_brewer(palette = "Set1") +
  scale_y_continuous(limits = c(-3, 3), breaks = -3:3) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  labs(title = "Combined log-ratio boxplot") + 
  mia_tcga_combat %>% 
  CPOP::boxplot_tbl(2) %>% 
  tidyr::separate(col = "object", into = c("sample_ID", "platform"), 
                  sep = "_", remove = FALSE) %>% 
  ggplot(aes(x = object, y = means,
             ymin = q1, ymax = q3,
             colour = platform)) +
  geom_point() +
  geom_linerange() +
  scale_colour_brewer(palette = "Set1") +
  scale_y_continuous(limits = c(-3, 3), breaks = -3:3) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  labs(title = "Combat boxplot") +
  patchwork::plot_layout(guides = "collect")
```

## Model generation

```{r}
set.seed(202201)

ns_combat_lasso = glmnet::cv.glmnet(
  x = ns_combat,
  y = list_binary_lassoy$ns,
  family = "binomial", alpha = 0.1, nfolds = 5)

ns_combat_lasso_pred_tcga = tibble(
    samples = rownames(list_lratio_binary$tcga),
    pred_prob = predict(ns_combat_lasso, 
                        newx = list_lratio_binary$tcga, 
                        type = "response", s = "lambda.min") %>% as.vector(),
    pred_class = predict(ns_combat_lasso, 
                         newx = list_lratio_binary$tcga, 
                         type = "class", s = "lambda.min") %>% as.vector()) %>%
  join_tcga_sample_data()

array_combat_lasso = glmnet::cv.glmnet(
  x = array_combat,
  y = list_binary_lassoy$array,
  family = "binomial", alpha = 0.1, nfolds = 5)

array_combat_lasso_pred_tcga = tibble(
    samples = rownames(list_lratio_binary$tcga),
    pred_prob = predict(array_combat_lasso, 
                        newx = list_lratio_binary$tcga, 
                        type = "response", s = "lambda.min") %>% as.vector(),
    pred_class = predict(array_combat_lasso, 
                         newx = list_lratio_binary$tcga, 
                         type = "class", s = "lambda.min") %>% as.vector()) %>%
  join_tcga_sample_data()
```

## Compare feature selection
```{r}
ns_combat_lasso_coef = coef(ns_combat_lasso, s = "lambda.min") %>% 
  as.matrix() %>% as.data.frame() %>% magrittr::set_colnames("coef_ns") %>% 
  tibble::rownames_to_column("lratio_name") %>% 
  tibble::as_tibble()
array_combat_lasso_coef = coef(array_combat_lasso, s = "lambda.min") %>% 
  as.matrix() %>% as.data.frame() %>% magrittr::set_colnames("coef_array") %>% 
  tibble::rownames_to_column("lratio_name") %>% 
  tibble::as_tibble()

combine_ns_array_combat_coef = inner_join(
  ns_combat_lasso_coef, 
  array_combat_lasso_coef,
  by = c("lratio_name"))

sum(combine_ns_array_combat_coef$coef_ns != 0)
sum(combine_ns_array_combat_coef$coef_array != 0)

combine_ns_array_combat_coef %>%
  group_by(sign_ns = sign(coef_ns), sign_array = sign(coef_array)) %>%
  tally() %>%
  ungroup() %>%
  tidyr::complete(sign_ns, sign_array, fill = list(n = 0)) %>%
  ggplot(aes(x = sign_ns, y = sign_array,
             label = n)) +
  geom_tile(colour = "grey20", fill = "grey80") +
  geom_text(size = 5) +
  scale_fill_distiller(palette = "Blues", direction = 1) +
  labs(x = "Sign of MIA-NanoString (Combat norm.) Lasso model coef.",
       y = "Sign of MIA-Microarray (Combat norm.) Lasso model coef.")
```

## Prediction values 

```{r}
combine_combat_pred_tcga = inner_join(
    x = ns_combat_lasso_pred_tcga %>% dplyr::select(samples, pred_prob, pred_class),
    y = array_combat_lasso_pred_tcga %>% dplyr::select(samples, pred_prob, pred_class),
    by = "samples", suffix = c("_ns", "_array"))

combine_combat_pred_tcga %>%
  ggplot(aes(x = pred_prob_ns, y = pred_prob_array)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(colour = "red") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1))

cat("Correlation: ", cor(combine_combat_pred_tcga$pred_prob_ns, 
                         combine_combat_pred_tcga$pred_prob_array), "\n")

cat("Concordance (yardstick): ", 
    yardstick::ccc_vec(combine_combat_pred_tcga$pred_prob_ns, 
                       combine_combat_pred_tcga$pred_prob_array), "\n")

cat("Ideality of correlation: ", 
    yardstick::iic_vec(combine_combat_pred_tcga$pred_prob_ns, 
                       combine_combat_pred_tcga$pred_prob_array), "\n")
```

## Survival prediction

### NanoString

```{r}
ns_combat_lasso_pred_tcga_surv = tibble(
    samples = rownames(list_lratio_survival$tcga),
    pred_prob = predict(ns_combat_lasso, 
                        newx = list_lratio_survival$tcga, 
                        type = "response", s = "lambda.min") %>% as.vector(),
    pred_class = predict(ns_combat_lasso, 
                         newx = list_lratio_survival$tcga, 
                         type = "class", s = "lambda.min") %>% as.vector()) %>%
  join_tcga_sample_data()

eval_prediction_df(ns_combat_lasso_pred_tcga_surv)
```

### Array

```{r}
array_combat_lasso_pred_tcga_surv = tibble(
    samples = rownames(list_lratio_survival$tcga),
    pred_prob = predict(array_combat_lasso, 
                        newx = list_lratio_survival$tcga, 
                        type = "response", s = "lambda.min") %>% as.vector(),
    pred_class = predict(array_combat_lasso,
                         newx = list_lratio_survival$tcga, 
                         type = "class", s = "lambda.min") %>% as.vector()) %>%
  join_tcga_sample_data()

eval_prediction_df(array_combat_lasso_pred_tcga_surv)
```


### NanoString + Array

```{r}
set.seed(2022)

ns_array_combat_lasso = glmnet::cv.glmnet(
  x = rbind(ns_combat, array_combat),
  y = c(list_binary_lassoy$ns, list_binary_lassoy$array),
  family = "binomial",
  alpha = 1,
  nfolds = 5)

ns_array_combat_lasso_pred_tcga_surv = tibble(
    samples = rownames(list_lratio_survival$tcga),
    pred_prob = predict(ns_array_combat_lasso,
                        newx = list_lratio_survival$tcga,
                        type = "response", s = "lambda.min") %>% as.vector(),
    pred_class = predict(ns_array_combat_lasso,
                         newx = list_lratio_survival$tcga,
                         type = "class", s = "lambda.min") %>% as.vector()) %>%
  join_tcga_sample_data()

eval_prediction_df(ns_array_combat_lasso_pred_tcga_surv)
```

# Bootstrap CPOP model, correlation and concodance

We will perform bootstrap on the columns of the melanoma data MIA-NanoString and MIA-Microarray to introduce perturbation in the CPOP model. We will do the same procedure on the Combat-MIA-NanoString-normalised and Combat-MIA-Microarray-normalised data. 

```{r, eval = FALSE}
# set.seed(2022)
# nexp = 100
# library(furrr)
# plan(multisession, workers = 5)
#
# boot_models_tbl = furrr::future_map(
#   .x = 1:nexp,
#   .f = ~ {
#     ns_index = sample(seq_len(nrow(list_lratio_binary$ns)), replace = TRUE)
#     array_index = sample(seq_len(nrow(list_lratio_binary$ns)), replace = TRUE)
#
#     mia_cpop_binary = CPOP::cpop_model(
#       z1 = list_lratio_binary$ns[ns_index, ],
#       z2 = list_lratio_binary$array[array_index, ],
#       y1 = list_binary_lassoy$ns[ns_index],
#       y2 = list_binary_lassoy$array[array_index],
#       family = "binomial",
#       alpha = 0.1, folds = 5,
#       n_features = 20)
#
#     ns_combat_lasso = glmnet::cv.glmnet(
#       x = ns_combat[ns_index, ],
#       y = list_binary_lassoy$ns[ns_index],
#       family = "binomial", alpha = 0.1,
#       nfolds = 5)
#
#     array_combat_lasso = glmnet::cv.glmnet(
#       x = array_combat[array_index,],
#       y = list_binary_lassoy$array[array_index],
#       family = "binomial", alpha = 0.1,
#       nfolds = 5)
#
#     return(list(mia_cpop_binary = mia_cpop_binary,
#                 ns_combat_lasso = ns_combat_lasso,
#                 array_combat_lasso = array_combat_lasso))
#   }, .progress = TRUE)
#
# boot_models_tbl = boot_models_tbl %>% magrittr::set_names(value = paste0("exp_", 1:nexp))
# saveRDS(boot_models_tbl, file = "../data/processed_data/boot_models_raw_and_combat_tbl_23_Jan_2022.rds")
```

```{r}
boot_models_tbl = readRDS(file = "../data/processed_data/boot_models_raw_and_combat_tbl_23_Jan_2022.rds")
```


```{r}
cpop_boot_pred_tcga_tbl = boot_models_tbl %>%
  purrr::map("mia_cpop_binary") %>%
  purrr::map_dfr(
    .x = .,
    .f = ~ predict_cpop(.x, newz = list_lratio_binary$tcga),
    .id = "exp_num") %>%
  join_tcga_sample_data() %>%
  dplyr::rename(
    pred_prob = cpop_model_avg_prob,
    pred_class = cpop_model_avg_class)

cpop_boot_pred_tcga_tbl %>%
  ggplot(aes(
    x = CPOP::expit(cpop_model1),
    y = CPOP::expit(cpop_model2),
    group = exp_num)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_abline(colour = "red") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1))

cpop_boot_concord_tbl = cpop_boot_pred_tcga_tbl %>%
  group_by(exp_num) %>%
  summarise(
    corr = cor(CPOP::expit(cpop_model1), CPOP::expit(cpop_model2)),
    ccc = yardstick::ccc_vec(CPOP::expit(cpop_model1), CPOP::expit(cpop_model2)))
```

## Lasso evaluation

```{r}
lasso_boot_pred_tcga_tbl = boot_models_tbl %>%
  purrr::map_dfr(.f = ~{
    ns_combat_lasso_pred_tcga = tibble(
      samples = rownames(tcga_combat),
      pred_prob = predict(.x$ns_combat_lasso,
                          newx = tcga_combat,
                          type = "response", s = "lambda.min") %>% as.vector(),
      pred_class = predict(.x$ns_combat_lasso,
                           newx = tcga_combat,
                           type = "class", s = "lambda.min") %>% as.vector())

    array_combat_lasso_pred_tcga = tibble(
      samples = rownames(tcga_combat),
      pred_prob = predict(.x$array_combat_lasso,
                          newx = tcga_combat,
                          type = "response", s = "lambda.min") %>% as.vector(),
      pred_class = predict(.x$array_combat_lasso,
                           newx = tcga_combat,
                           type = "class", s = "lambda.min") %>% as.vector())

    combine_combat_pred_tcga = inner_join(
      x = ns_combat_lasso_pred_tcga %>% dplyr::select(samples, pred_prob, pred_class),
      y = array_combat_lasso_pred_tcga %>% dplyr::select(samples, pred_prob, pred_class),
      by = "samples", suffix = c("_ns", "_array"))
    return(combine_combat_pred_tcga)
    }, .id = "exp_num") %>%
  dplyr::mutate(samples = samples %>% str_remove_all("_tcga")) %>%
  join_tcga_sample_data()


lasso_boot_pred_tcga_tbl %>%
  ggplot(aes(
    x = pred_prob_ns,
    y = pred_prob_array,
    group = exp_num)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  geom_abline(colour = "red") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1))

lasso_boot_concord_tbl = lasso_boot_pred_tcga_tbl %>%
  group_by(exp_num) %>%
  summarise(
    corr = cor(pred_prob_ns, pred_prob_array),
    ccc = yardstick::ccc_vec(pred_prob_ns, pred_prob_array))
```

## Combing concordance metrics

```{r, fig.height=7}
combine_boot_concord_tbl = inner_join(
  cpop_boot_concord_tbl, lasso_boot_concord_tbl, by = "exp_num",
  suffix = c("_cpop", "_lasso"))

combine_boot_concord_tbl_long = combine_boot_concord_tbl %>%
  pivot_longer(cols = -exp_num) %>%
  tidyr::separate(col = name, into = c("metric", "model"), remove = FALSE) %>%
  dplyr::mutate(metric = case_when(
    metric == "ccc" ~ "Concordance",
    metric == "corr" ~ "Correlation",
    TRUE ~ NA_character_))

combine_boot_concord_tbl_long %>%
  ggplot(aes(x = model, y = value)) +
  geom_boxplot() +
  facet_wrap(~metric, scales = "free_y")

combine_boot_concord_tbl_long_ccc = combine_boot_concord_tbl_long %>%
  dplyr::filter(metric == "Concordance")

pairwise.wilcox.test(x = combine_boot_concord_tbl_long$value, g = combine_boot_concord_tbl_long$name)
```


